Passenger Prediction in Shared Accounts for Flight Service Recommendation

### 2016-07-10 Paper.tex
Add files via upload

\section{Introduction}
\label{sec:intro}
Recent years, with the rapid development of online travel agencies, there are more and more passengers booking flights through online travel service providers. Typically, an user may input his departure city , arrival city  and departure time, then he will get a list of candidate flights. In general, there are dozens of flights within a list .  Figure \ref{fig:res} shows a list for all candidate flights  from Beijing to Hong Kong when input a query. There are total 156 matchable flights and the website only provides simple sorting strategies such as departure time, arrival time or price. Thus, it may take quite a long time for users to find the appropriate one. Therefore, a recommender system is  necessary for better user experiences.
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.48]{flight_tickets.eps}
\caption{Search result for flight tickets}
\label{fig:res}
\end{figure}\par
Essentially, a recommender system is used for generating a personalized ranking on a set of items\cite{lv:rec}. Typical recommender systems assume that the preference model of an individual can be learned from each account. However, for online flight booking, it is quite common that a user books flights for his family members, friends or colleagues.  Obviously, every passenger has his own preferences, which should be considered in flight recommendation. Unfortunately, before placing the order, people will not provider passengers' information.  Therefore, if we want to improve the accuracy of recommendation, we have to predict who are going to fly.  \par
 To predict who is using the account to place orders is regarded as the shared account problem. Some approaches have been proposed to extract implicit user information within an account. For example,  in \cite{kabbur:nlmf}\cite{jason:embedding} the nonlinear latent factorization is applied to  extract user information. However, these methods only rely on the analysis of relationships between products and users \cite{koen:top-n}\cite{yutaka:modeling}, and context information (such as time information, the query input) that can help differentiate users is neglected. Besides, current methods always assume an account represents a fixed number of implicit users, which are not always true in reality. Moreover, so far, the experiments have only been evaluated on artificial composite datasets. 
 The flight booking service providers can acquire passenger's personal information after he has submitted the order. Thus, our goal is to predict distributions of passengers in current session based on historical behaviors and contextual information so that we can make flight recommendations in a more precise way. To achieve this target, we adopt the probabilistic author-topic model to analyze passengers' behaviors. We regard all historical orders in an account as the corpus, all passengers appearing in the account as authors and orders together with context information as documents. Note that if an order has more than one passenger, they are regarded as co-authors of this document.\par
 In experiment section, we apply our method to a real flight order dataset.  And the result shows that our method has higher identification accuracy and a better recommendation accuracy.\par
 This paper has following contributions:\\
\begin{enumerate}
\item We provide a probabilistic passenger identification method for online flight booking service.
\item We apply the passenger prediction results to flight recommendation.
\item We evaluate our approach on a real flight order dataset, and the result shows that our approach can effectively improve the recommendation accuracy.
\end{enumerate}

@@

\section{Related Work}\par
\label{sec:relw}
\textbf{Content based Recommendation.} In online travel service industry such as airline tickets or hotel booking, the item (for example, an air ticket, a hotel booking order) information is usually highly structured, however, this information is quite mutable. Content-based recommender systems \cite{tech:cset}\cite{lops:handbook}are appropriate for highly structured items. The method needs proper representations of the attributes of items and the profile of user interests, then it tries to match  the profile of user's interests with attributes of items. In general, a content-based recommender system consists of three components. A content analyzer extracts structured attributes of items through a series of pre-processing steps and it produces the inputs for the system, which always needs domain related knowledge. A profile learner analyzes and constructs users' profiles of interests by collecting users' historical behaviors. In addition, this component also relies on feedbacks of users. A filtering component evaluates the relevances of items to particular users by matching them with users' profiles and the outputs can be either binary or continuous. In this stage, some similarity measurements including cosine or euclidean distance can usually be applied.\par
\textbf{Flight Ticket Recommendation.} There is little work focusing on personalized flight  recommendations. Many online flight booking agencies provide sorting and filtering strategies to improve users' experiences, such as sorting by price or takeoff time and filtering by airline or departure city. Others also provide additional information that can help users, for example, placing the most popular flights at the top or highlighting some of them. These methods do make sense in helping users find appropriate flights within less time. However, they can not provide personalized recommendations. In 2004, Lorcan Coyle\cite{lor:flight} proposed a Personal Travel Assistant(PTA). PTA is based on case-based reasoning, which takes users' historical orders as cases and recommends the most similar candidate flights to users' queries. In addition, PTA also proposes the attribute weight training and case storage method.\par
\textbf{Implicit Feedback.} The users' feedbacks can be divided into implicit or explicit type. The explicit feedbacks can be acquired from users' ratings or text comments. However,  users may not willing to provide explicit feedbacks since it needs additional efforts. Moreover, in many applications, a single numeric rating may not be adequate for describing users feeling to an item. Implicit feedback methods\cite{song:mining} try to infer users' attitudes to an item based on their behaviors, such as clicking, collecting or bookmarking et al. It has the advantage of being free from the problem of lacking feedback data. Nevertheless, implicit feedback based recommender suffers from problems of lack of negative feedbacks and data noises. In order to provide recommendations when only implicit feedbacks are available, some methods are proposed, such as one class collaborative filtering and content based recommender.\par

\textbf{Probabilistic Topic Models.}  Topic models\cite{blei:lda}\cite{mic:atm}\cite{mark:prob} are proposed for automated extraction of useful semantic information from corpus data, which are widely used in information retrieval, document annotation, etc. The main step is extracting latent factors from the corpus, named topics, which are commonly probabilistic distributions over words. Topic models have three major advantages over other document modeling methods. At first, topic models provide a completely unsupervised approach to extract topics, thus requiring no document labels and no initialization. Unsupervised methods are necessary for modeling large document corpus, besides, in some fields, with the rapid change of content's topics, predefined topic categories may not reflect the development of specific field. Secondly, latent topics are interpretable respectively, so that extracted results are understandable for users. Finally, each document may consist of multiple topics and it can be considered as probabilistic distributions over topics. Many existing models regard documents as mixtures of topics, such as pLSI and LDA. The author-topic model is a generative model that extends LDA to include authorship information for document modeling. In our approach, the topic model is innovatively applied to identify users. \par

@@

\section{Passenger Prediction Model}
\label{sec:mod}
In this Section, we propose a generic model for predicting the probabilistic distribution of passengers, which can be used for recommendation.  Figure 2 shows an overview of the recommendation process. At first, a dataset of orders is saved into the database, in which each account may correspond to several passengers. We use this dataset to train the passenger prediction model as well as to extract passengers' preferences. Before a recommendation process starts, we calculate passenger distribution probability applying trained model and contextual information of current session. At the final step, we generate a ranked candidate list as the recommendation result based on passenger prediction and passengers' preference information.\par
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.40]{overview.eps}
\caption{Overview of flight recommendation based on passenger prediction}
\label{fig:over}
\end{figure}\par

\subsection{Model description and notation}
Topic models are widely used in recommender systems. For example, in pLSA model\cite{tomas:coll}, a topic is a multinomial distribution over items and it represents a latent feature. An account is a multinomial distribution over topics and it can represent the preferences of an account. Each purchase can be regarded a sample that selects a topic $z$ from account $a$, and takes an item from topic $z$.\par
When the topic model is applied to flight recommendation, there are some challenges. Firstly, the concepts behind flight tickets are quite difficult to be modeled. Conventional items only have static features, which means that the contents of an item are less likely to change in the future. For flight tickets, however, the price changes frequently even for the same flight number and class. Since the price factor plays a significant role in users' choice, so we can't consider the tickets with different prices as different items. Secondly, the amount of tickets for a flight is limited, which leads to a very sparse user-item matrix so that we can't use a collaborative-filtering like method directly.\par
Fortunately, the decisive factors of a flight ticket are quite fixed, such as airline, takeoff time, price and class etc. Some factors are discrete while others are continuous. For continuous factors, we can partition them into discrete non-overlapping intervals. Thus, the number of discrete alternative values of each factor is also very limited. We take orders within each account as a corpus. Our notations are summarized in table \ref{tab:not}.\par
\begin{table}[!htbp]
\centering
\caption{\label{tab:not}Notation}
\begin{tabular}{|c|c|} \hline
$M$ & number of accounts within the dataset\\ \hline
$V$ & number of words in vocabulary\\ \hline
$O$ & orders in an account\\ \hline
$P$ & passengers in an account\\ \hline
$P_i$ & passengers for order $O_i$ \\ \hline
$F$ & selected factors\\ \hline
$V$ & size of vocabulary\\ \hline
$K$ & number of topics\\ \hline
\end{tabular}
\end{table}

In author-topic model\cite{deiv:corpora}, a predefined vocabulary is generated containing discrete alternatives and intervals of all selected factors, we use integers to donate each entry in the vocabulary. The model discovers both topics distributed in an order and passengers associated with the distribution of topics. We treat every order as a bag of words, thus deduce each order to a vector of word counts. An order can be considered as a document with fixed word length, which can be represented as a vector of integers. All passengers corresponding to an order are co-authors for that document. Each passenger is associated with a multinomial distribution over topics and each topic is associated with a multinomial distribution over entries. The model essentially reduces dimensions of documents, depending on the amount of topics.\par

For an account $M$, we generate the observed passenger list $P$. We denote authors' distribution over topics by a $|P| \times K$ matrix $\Theta$. The multinomial distribution can be generated from Dirichlet prior distribution with hyper-parameter $\alpha$. Topics' distribution over words is denoted by a $K \times \sum_{i=1}^{|F|}|F_i| $  matrix $\Phi$, also, distribution can be generated with hyper-parameter $\beta$. Generally, these hyper-parameters needn't be estimated, here, we fix $\alpha$ and $\beta$ at 50/K and 0.01 respectively. \par

@@

\begin{enumerate}
\item For each passenger $p=1,\dots,|P|$ draw $\Theta_p \sim Dirichlet(\alpha)$
\item For each topic $t=1,\dots,K$ draw $\Phi_t \sim Dirichlet(\beta)$
\item For each order $o=1,\dots,O$
       \begin{enumerate}[fullwidth,itemindent=1em,label=(\alph*)]
       \item given passengers $P$
       \item For each word $i=1,\dots, N_o$
              \begin{enumerate}[fullwidth,itemindent=2em,label=(\roman*)]
              \item draw a passenger $X_{oi} \sim Uniform(P)$
              \item draw a topic $Z_{oi} \sim Discrete(\theta_{X_{oi}})$
              \item draw a word $w_{oi} \sim Discrete(\phi_{Z_{oi}})$
              \end{enumerate}
       \end{enumerate}
\end{enumerate}
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.55]{f1.eps}
\caption{Probability Graph for the Author-Topic Model}
\label{fig:mod}
\end{figure}\par
Figure \ref{fig:mod} shows the graphical model corresponding to this process, where shaded and non-shaded nodes represent observed and unobserved variables respectively. It's a plate notation, the number at the bottom of each box indicates repeated times of the operation. Arrows represents the conditional dependency between variables. In the author-topic model, observed variables include words in an order and the passengers of the order.

\subsection{Parameter Inference}
As mentioned above, the author-topic model includes two sets of unknown parameters, the $P$ passenger-topic distributions $\theta$ and the $K$ topic-word distributions $\phi$. There are a variety of algorithms that can be applied to estimate the parameters of author-topic models, the Expectation-Maximization algorithm and Gibbs sampling\cite{gregor:esti}. Generic EM algorithms are likely to face local maximum and computational inefficiency problems. Here we utilize Gibbs sampling because it does not explicitly estimate parameters, instead, it evaluates posteriori distribution just based on drawn passengers $X$ and topics $Z$. Thus it is simple for Dirichlet priors.\par
We can obtain the probability of every word $\mathbf{w}_o$ generated in each order, conditioned on $\Theta$ and $\Phi$ is:\\
\begin{flalign}
\begin{split}
\label{eq:1} 
& P(\mathbf{w}_o | \Theta,\Phi,P) =\prod_{i=1}^{N_o}P(w_{oi}|\Theta,\Phi,\mathbf{p}_o) \\
& =\prod_{i=1}^{N_o}\sum_{p=1}^{|P|}\sum_{t=1}^{K}P(w_{oi}|z_{oi}=t,\Phi)
P(z_{oi}=t|x_{oi}=p,\Theta)P(x_{oi}=p|\mathbf{p}_o)\\
& =\prod_{i=1}^{N_o}\frac{1}{|P|}\sum_{p \in p_o}\sum_{t=1}^{K}\phi_{w_{oi}t}\theta_{tp}
\end{split} &
\end{flalign}

@@

The algorithm performs in three steps, i.e., initialization, sampling and updating respectively. At the first step, we assign each word in the corpus with random passengers and topics. Then for every topic $t$, we count the frequency of each vocabulary item assigned with $t$. For every passenger $p$, we count the frequency of each $K$ topic assigned with $p$. For every sampling operation, a word in corpus is chosen, the probability of topic distribution and passenger distribution is calculated conditioned on the rest words in corpus by applying Equation \ref{eq:2}, with these two probabilities, we can sample a new topic and passenger for the current word. After several batch iterations, the passenger-topic matrix $\Theta$ and topic-word matrix $\Phi$ can be updated by applying Equation \ref{eq:3}, \ref{eq:4}. So the computational efficiency is the number of words multiplies number of topics, passengers and iteration times.\par

@@

One important issue here is how to determine the number of passengers. Through statistic analysis we find that most of testing orders have exactly one passenger. On average, the probability of every passenger is $P_A = \frac{1}{|P|}$. We determine the passenger number by omitting passengers whose probability is not greater than $P_A$. We show the evaluation metric and prediction result in Section 4.3.\par

@@

\begin{algorithm}[htb]
\caption{trainATM}
\begin{algorithmic}[1]
\label{alg:2}
\REQUIRE
Word vector $\overrightarrow{W}$. \par
A test anonymous order $to$
\ENSURE 
User-topic parameter $\Theta$ \par
Topic-word parameter $\Phi$
\STATE $nak , nkv, nak\_sum, nkv\_sum \leftarrow \mathbf{0}$
\FORALL{List o : $\overrightarrow{W}$}
\FORALL{Word w : o}
\STATE pick a topic $t$ and a user $p$ Multinomial;
\STATE user-topic count $nak[p][t] += 1$;
\STATE user-topic sum $nak\_sum[p] += 1$;
\STATE topic-word count $nkv[t][w] += 1$;
\STATE topic-word sum $nkv\_sum[t] += 1$;
\ENDFOR
\ENDFOR
\WHILE {not reach iteration limit}
\FORALL{List o : $\overrightarrow{W}$}
\FORALL{Word w : o}
\STATE $nak[p][t] -= 1, nak\_sum[p] -= 1$;
\STATE $nkv[t][w] -= 1, nkv\_sum[t] -= 1$;
\STATE sample a new user $p'$ and topic $t'$ by Equation \ref{eq:2};
\STATE $nak[p'][t'] += 1, nak\_sum[p'] += 1$;
\STATE $nkv[t'][w] += 1, nkv\_sum[t'] += 1$;
\ENDFOR
\ENDFOR
\ENDWHILE
\STATE calculate $\Theta,\Phi$ by Equation \ref{eq:3} ,\ref{eq:4};
\RETURN $\Theta,\Phi$;
\end{algorithmic} 
\end{algorithm}

@@

\subsection{Factor selection and frequency statistic}
A flight ticket contains a set of attributes such as price, takeoff time, craft size, etc. We should select some of them as the decisive factors. These attributes may be either discrete or continuous variables. We divide a continuous attribute into a vector of intervals, so any factor can be represented as a vector of finite alternatives. Then, we analyze user's submitted orders by counting the frequency of all alternatives for each factor. In a word, a user's preference model can be described as a union of vectors, a vector indicates one factor and an element within one vector represents the frequency of corresponding alternative which the user has chosen.\par
We can also represent a candidate flight ticket as a vector, where each element is the corresponding alternative of each factor. So, for every ticket in candidate list, we compare the vector with users' preference model and get a score for every factor based on the frequency. We can sum up score of every factor to get a total score for that item, and rank all candidates according to the total score. We can then generate a top-$N$ recommendation.\par
\subsection{Weight of factors}
Since users may care different factors, we introduce personalized weight vector for each user. The evaluation of this weight vector contains two steps, initializing and training. At the initializing step, we estimate how much a user focuses on a factor through the information entropy, as mentioned in \cite{yang:pers}. We believe that if a user focuses on a factor, his behavior will be more concentrated. The entropy $H$ is a measure to describe the uncertainty for discrete random variable $X$ and probability mass function $P(X)$. Here is the mathematical representation for entropy.\\
\begin{equation}
H(X)=E[-lnP(X)]= - \sum_{i-1}^n P(x_i)log_bP(X_i)
\end{equation}\par
The value $b$ is the base of logarithm, and is usually 2. The entropy of every factor can be calculated and initialized as the weight value after normalization. Considering that different factors have a variety quantity of alternatives and the user may concern different factors some time later, we update the weight vector once the user submits an order. For each submission, we have already produced a ranked list for recommendations, if the user doesn't choose an item (i.e, a ticket) from top-$N$ items, we consider this recommendation fails, so that the weight vector need be retrained to make the rank of selected item improved. The concept of weight training is mentioned in \cite{lor:flight}. By performing weight training for factors, we hope that the result of recommendation getting better in the future. The training procedure is quite straightforward. At first a learning rate should be set, such as $\alpha = 0.1$ . If the selected item has lower similarity for factor $f$ than items ranked before it, the weight value of this factor should be decreased. Otherwise, the weight value should be increased. To avoid the problem of over-fitting, we define a maximum number of iteration step. So the stop criteria is either $1)$ the rank of selected items can't be improved anymore or $2)$ the procedure reaches the iteration time limitation.\\
The preference based recommendation algorithm is described as Algorithm \ref{alg:3}. First extracting a user's preference  $P$ and initializing weighted list $W$ through entropy. Items in candidate list are ranked based on inner product of $W$ and corresponding value of factor. After every recommendation, the list is retrained and stored depending on the item that user actually chosen.

\begin{algorithm}[htb]
\caption{Preference-Based-Recommendation}
\begin{algorithmic}[1]
\label{alg:3}
\REQUIRE
User's ID $u$ \par
User's history order list $O$ \par
Candidate item list $C$
\ENSURE Ranked candidate list $R$
\STATE Define a factor list $F$;
\STATE $P \leftarrow extractUserPref(O,F)$;
\STATE $W \leftarrow fetchWeightedList(u)$;
\IF{$W$ is $null$}
\STATE initialize $W$ by entropy;
\ENDIF
\STATE $R \leftarrow \emptyset$;
\FORALL { $c : C$} 
\STATE append $(c,weightedScore(c,W))$ to R;
\ENDFOR 
\STATE sort $R$ by score;
\STATE $trainWeightedList(W,R,P)$;
\STATE $saveWeightedList(u,W)$;
\RETURN $R$;
\end{algorithmic} 
\end{algorithm}

\begin{algorithm}[htb]
\caption{extractUserPref}
\begin{algorithmic}[1]
\label{alg:4}
\REQUIRE
User's history order list $O$. \par
Predefined discrete factor list $F$
\ENSURE 
User's preference list $P$
\STATE List $P \leftarrow \emptyset$;
\FORALL {List $f : F$} 
\STATE append  $List(0,|f|)$  to $P$;
\ENDFOR
\FORALL{List $order : O$}
\FOR{$i: 1 \to |order| $}
\STATE $P[i][order_i] += 1$;
\ENDFOR	
\ENDFOR 
\RETURN $P$;
\end{algorithmic} 
\end{algorithm}

\begin{algorithm}[htb]
\caption{weightedScore}
\begin{algorithmic}[1]
\label{alg:5}
\REQUIRE \par
Candidate item $c$ \par
User's preference list $P$ \par
User's weight vector $W$ 
\ENSURE Score of candidate $S$
\STATE $S \leftarrow 0$;
\FOR{$i: 1 \to |c| $}
\STATE $S += \frac{P[i][c_i]}{sum(P[i])} \times W[i]$;
\ENDFOR
\RETURN $S$;
\end{algorithmic} 
\end{algorithm}

\begin{algorithm}[htb]
\caption{trainWeightedList}
\begin{algorithmic}[1]
\label{alg:6}
\REQUIRE \par
Ranked list $R$ \par
User's Preference list $P$ \par
User's weight list $W$ \par
User actually chose candidate $c$
\ENSURE Trained weighted list $W$
\STATE $R_c \leftarrow rankOf(R,c)$;
\STATE $\alpha \leftarrow 0.1$;
\STATE $count \leftarrow 0$;
\WHILE {$count < limit$}
\STATE $BeatList \leftarrow List(false,|W|)$;
\FOR{$i: 1 \to |W|$}
\FOR{$j: 1 \to R_c $}
\IF{$P[i][R_{ji}] > P[i][c_{i}] $}
\STATE $Beatlist[i] \leftarrow true$;
\STATE break;
\ENDIF
\ENDFOR
\IF{$BeatList[i] = true$}
\STATE $BeatList[i] \leftarrow false$;
\STATE $W[i] \leftarrow W[i] / (1+\alpha)$;
\ENDIF
\ENDFOR
\ENDWHILE
\STATE $normalize(W)$;
\RETURN $S$;
\end{algorithmic} 
\end{algorithm}

\section{Experiments}
\label{sec:exp}
In the following subsections, we introduce datasets and evaluation metrics. Then we evaluate the performance of passenger prediction and passenger prediction based recommendation respectively.
\subsection{Dataset}
In our experiments, we use a real flight ticket dataset, consisting of all submitted orders for two years. An order contains account id and all passengers' id. We use desensitized passenger id to distinguish individuals from each other. Besides, an order contains some travel information that can be considered as decisive factors, such as departure airport, arrival airport, takeoff time, etc. In our experiment, we select active users whose amount of historical orders reaches a threshold, thus passengers' preferences can be extracted with more confidence. In addition, many researches\cite{amy:guess}\cite{yutaka:modeling} on shared account recommendation generate an artificial dataset by composing single accounts' data. We also generate an artificial dataset combining single passengers' data from different accounts. The overviews of these two datasets and the meta data of orders are summarized in table \ref{tab:dat} and table \ref{tab:mdat}.
\begin{table}[!htbp]
\centering
\caption{\label{tab:dat}Datasets}
\begin{tabular}{|c|c|c|c|} \hline
& \# accounts & \# passengers & \# orders\\ \hline
Real & 4632&7034&38907 \\ \hline
Artificial &1604&3208&29759\\ \hline
\end{tabular}
\end{table}

>>>>

\section{Introduction}
\label{sec:intro}
Recent years, with the rapid development of online travel agencies, there are more and more passengers booking flights through online travel service providers. Typically, an user may input his departure city, arrival city and departure time, then he will get a list of candidate flights. In general, there are dozens of flights within a list. We perform a search for flights from Beijing to Hong Kong at a website, there are total 156 matchable flights. The website only provides simple sorting strategies such as departure time, arrival time or price. Thus, it may take quite a long time for users to find the appropriate one. Therefore, a recommender system is necessary for better user experiences.

\par Essentially, a recommender system is used for generating a personalized ranking on a set of items\cite{lv:rec}. Typical recommender systems assume that the preference model of an individual can be learned from each account. However, for online flight booking, it is quite common that a user books flights for his family members, friends or colleagues.  Obviously, every passenger has his own preferences, which should be considered in flight recommendation. Unfortunately, before placing the order, people will not provider passengers' information.  Therefore, if we want to improve the accuracy of recommendation, we have to predict who are going to fly.  \par
To predict who is using the account to place orders is regarded as the shared account problem. Some approaches have been proposed to extract implicit user information within an account. For example,  in \cite{kabbur:nlmf}\cite{jason:embedding} the nonlinear latent factorization is applied to  extract user information. However, these methods only rely on the analysis of relationships between products and users \cite{koen:top-n}\cite{yutaka:modeling}, and context information (such as time information, the query input) that can help differentiate users is neglected. Besides, current methods always assume an account represents a fixed number of implicit users, which are not always true in reality. Moreover, so far, the experiments have only been evaluated on artificial composite datasets. \par
The flight booking service providers can acquire passenger's personal information after he has submitted the order. Thus, our goal is to predict distributions of passengers in current session based on historical behaviors and contextual information so that we can make flight recommendations in a more precise way. To achieve this target, we adopt the probabilistic author-topic model to analyze passengers' behaviors. We regard all historical orders in an account as the corpus, all passengers appearing in the account as authors and orders together with context information as documents.\par
In experiment section, we apply our method to a real flight order dataset. And the result shows that our method has higher identification accuracy and a better recommendation accuracy.\par
This paper has following contributions:\\
\begin{enumerate}
\item We provide a probabilistic passenger identification method for online flight booking service.
\item We apply the passenger prediction results to flight recommendation.
\item We evaluate our approach on a real flight order dataset, and the result shows that our approach can effectively improve the recommendation accuracy.
\end{enumerate}

@@

\section{Related Work}\par
\label{sec:relw}
\textbf{Content based Recommendation.} In online travel service industry such as airline tickets or hotel booking, the item (for example, an air ticket, a hotel booking order) information is usually highly structured, however, this information is quite mutable. Content-based recommender systems \cite{tech:cset}\cite{lops:handbook}are appropriate for highly structured items. The method needs proper representations of the attributes of items and the profile of user interests, then it tries to match the profile of user's interests with attributes of items.\par
\textbf{Flight Ticket Recommendation.} There is little work focusing on personalized flight  recommendations. Many online flight booking agencies provide sorting and filtering strategies to improve users' experiences, such as sorting by price and filtering by airline etc. These methods do make sense in helping users find appropriate flights within less time. However, they can not provide personalized recommendations. Flight recommendation mainly bases on implicit feedback\cite{song:mining}, which suffers from problems of data noises and lack of negative feedbacks. In 2004, Lorcan Coyle\cite{lor:flight} proposed a Personal Travel Assistant(PTA). PTA is based on case-based reasoning, which takes users' historical orders as cases and recommends the most similar candidate flights to users' queries.\par

\textbf{Probabilistic Topic Models.}  Topic models\cite{blei:lda}\cite{mic:atm}\cite{mark:prob} are proposed for automated extraction of useful semantic information from corpus data. The main step is extracting latent factors from the corpus, named topics, which are commonly probabilistic distributions over words. Topic models provide a completely unsupervised approach to extract topics, thus requiring no document labels and no initialization. Besides, each document may consist of multiple topics and it can be considered as probabilistic distributions over topics. The author-topic model is a generative model that extends LDA to include authorship information for document modeling. In our approach, the topic model is innovatively applied to identify users. \par

@@

\section{Passenger Prediction Model}
\label{sec:mod}
In this Section, we propose a generic model for predicting the probabilistic distribution of passengers, which can be used for recommendation.  Figure \ref{fig:over} shows an overview of the recommendation process. At first, a dataset of orders is saved into the database, in which each account may correspond to several passengers. We use this dataset to train the passenger prediction model as well as to extract passengers' preferences. Before a recommendation process starts, we calculate passenger distribution probability applying trained model and contextual information of current session. At the final step, we generate a ranked candidate list as the recommendation result based on passenger prediction and passengers' preference information.\par
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.40]{overview.eps}
\caption{Overview of flight recommendation based on passenger prediction}
\label{fig:over}
\end{figure}\par

\subsection{Model description and notation}
Topic models are widely used in recommender systems. For example, in pLSA model\cite{tomas:coll}, a topic is a multinomial distribution over items and it represents a latent feature. An account is a multinomial distribution over topics and it can represent the preferences of an account. Each purchase can be regarded a sample that selects a topic $z$ from account $a$, and takes an item from topic $z$.\par
When the topic model is applied to flight recommendation, there are some challenges. Firstly, the concepts behind flight tickets are quite difficult to be modeled. Conventional items only have static features, which means that the contents of an item are less likely to change in the future. For flight tickets, however, the price changes frequently even for the same flight number and class. Since the price factor plays a significant role in users' choice, so we can't consider the tickets with different prices as different items. Secondly, the amount of tickets for a flight is limited, which leads to a very sparse user-item matrix so that we can't use a collaborative-filtering like method directly. Fortunately, the decisive factors of a flight ticket are quite fixed, such as airline, takeoff time, price and class etc. The number of alternatives for each factor is also quite limited.\par
Our notations are summarized in table \ref{tab:not}.\par
\begin{table}[!htbp]
\centering
\caption{\label{tab:not}Notation}
\begin{tabular}{|c|c|} \hline
$M$ & number of accounts within the dataset\\ \hline
$V$ & number of words in vocabulary\\ \hline
$O$ & orders in an account\\ \hline
$P$ & passengers in an account\\ \hline
$P_i$ & passengers for order $O_i$ \\ \hline
$F$ & selected factors\\ \hline
$V$ & size of vocabulary\\ \hline
$K$ & number of topics\\ \hline
\end{tabular}
\end{table}

In author-topic model\cite{deiv:corpora}, a predefined vocabulary is generated containing discrete alternatives and intervals of all selected factors. We treat every order as a bag of words, thus deduce each order to a vector of word counts. Each passenger is associated with a multinomial distribution over topics and each topic is associated with a multinomial distribution over entries.\par

For an account $M$, we generate the observed passenger list $P$. We denote passengers' distribution over topics by a $|P| \times K$ matrix $\Theta$. The multinomial distribution can be generated from Dirichlet prior distribution with hyper-parameter $\alpha$. Topics' distribution over words is denoted by a $K \times \sum_{i=1}^{|F|}|F_i| $  matrix $\Phi$, also, distribution can be generated with hyper-parameter $\beta$. Generally, these hyper-parameters needn't be estimated, here, we fix $\alpha$ and $\beta$ at 50/K and 0.01 respectively. \par

@@

\begin{enumerate}
\item For each passenger $p=1,\dots,|P|$ draw $\Theta_p \sim Dirichlet(\alpha)$
\item For each topic $t=1,\dots,K$ draw $\Phi_t \sim Dirichlet(\beta)$
\item For each order $o=1,\dots,O$
       \begin{enumerate}[fullwidth,itemindent=1em,label=(\alph*)]
       \item given passengers $P$
       \item For each word $i=1,\dots, N_o$
              \begin{enumerate}[fullwidth,itemindent=2em,label=(\roman*)]
              \item draw a passenger $X_{oi} \sim Uniform(P)$
              \item draw a topic $Z_{oi} \sim Discrete(\theta_{X_{oi}})$
              \item draw a word $w_{oi} \sim Discrete(\phi_{Z_{oi}})$
              \end{enumerate}
       \end{enumerate}
\end{enumerate}

\subsection{Parameter Inference}
As mentioned above, the author-topic model includes two sets of unknown parameters, the $P$ passenger-topic distributions $\theta$ and the $K$ topic-word distributions $\phi$. Generic EM algorithms are likely to face local maximum and computational inefficiency problems. Here we utilize Gibbs sampling\cite{gregor:esti} because it does not explicitly estimate parameters, instead, it evaluates posteriori distribution just based on drawn passengers $X$ and topics $Z$. Thus it is simple for Dirichlet priors.\par
We can obtain the probability of every word $\mathbf{w}_o$ generated in each order, conditioned on $\Theta$ and $\Phi$ is:\\
\begin{flalign}
\begin{split}
\label{eq:1} 
& P(\mathbf{w}_o | \Theta,\Phi,P) =\prod_{i=1}^{N_o}P(w_{oi}|\Theta,\Phi,\mathbf{p}_o) \\
& =\prod_{i=1}^{N_o}\sum_{p=1}^{|P|}\sum_{t=1}^{K}P(w_{oi}|z_{oi}=t,\Phi)
P(z_{oi}=t|x_{oi}=p,\Theta)P(x_{oi}=p|\mathbf{p}_o)\\
& =\prod_{i=1}^{N_o}\frac{1}{|P|}\sum_{p \in p_o}\sum_{t=1}^{K}\phi_{w_{oi}t}\theta_{tp}
\end{split} &
\end{flalign}

@@

The algorithm performs in three steps, i.e., initialization, sampling and updating respectively. At the first step, we assign each word in the corpus with random passengers and topics. For every sampling operation, a word in corpus is chosen, the probability of topic distribution and passenger distribution is calculated conditioned on the rest words in corpus by applying Equation \ref{eq:2}, with these two probabilities, we can sample a new topic and passenger for the current word. After several batch iterations, the passenger-topic matrix $\Theta$ and topic-word matrix $\Phi$ can be updated by applying Equation \ref{eq:3}, \ref{eq:4}. So the computational efficiency is the number of words multiplies number of topics, passengers and iteration times.\par

@@

@@

@@

\subsection{Factor selection and frequency statistic}
A flight ticket contains a set of attributes. These attributes may be either discrete or continuous variables. We divide continuous attribute into a list of intervals, for example, takeoff time can be divided into morning, noon, afternoon, night etc. Thus, any factor can be represented by finite alternatives. Then, we count the distribution of alternatives from a user's historical orders to analyze user's preference on corresponding factor. In a word, the preference model can be described as a union of vectors, a vector indicates one factor and an element within one vector represents the frequency of corresponding alternative which the user has chosen.\par
We can also represent a candidate flight ticket as a vector, where each element is the corresponding alternative of each factor. So, for every ticket in candidate list, we compare the vector with users' preference model and get a score for every factor based on the frequency. We can sum up score of every factor to get a total score for that item, and rank all candidates according to the total score. We can then generate a top-$N$ recommendation.\par
\subsection{Weight of factors}
Since users may care different factors, we introduce personalized weight vector for each user. We estimate how much a user focuses on a factor through the information entropy, as mentioned in \cite{yang:pers}. We believe that if a user focuses on a factor, his behavior will be more concentrated. The entropy $H$ is a measure to describe the uncertainty for discrete random variable $X$ and probability mass function $P(X)$. Here is the mathematical representation for entropy.\\
\begin{equation}
H(X)=E[-lnP(X)]= - \sum_{i-1}^n P(x_i)log_bP(X_i)
\end{equation}\par
The value $b$ is the base of logarithm, and is usually 2. The entropy of every factor can be calculated and initialized as the weight value after normalization. \par
The preference based recommendation algorithm is described as Algorithm \ref{alg:3}. First extracting a user's preference  $P$ and initializing weighted list $W$ through entropy. Items in candidate list are ranked based on inner product of $W$ and corresponding value of factor.

\begin{algorithm}[htb]
\caption{Preference-Based-Recommendation}
\begin{algorithmic}[1]
\label{alg:3}
\REQUIRE
User's ID $u$ \par
User's history order list $O$ \par
Candidate item list $C$
\ENSURE Ranked candidate list $R$
\STATE Define a factor list $F$;
\STATE $P \leftarrow extractUserPref(O,F)$;
\STATE initialize $W$ by entropy;
\STATE $R \leftarrow \emptyset$;
\FORALL { $c : C$} 
\STATE append $(c,weightedScore(c,W))$ to R;
\ENDFOR 
\STATE sort $R$ by score;
\RETURN $R$;
\end{algorithmic} 
\end{algorithm}

\section{Experiments}
\label{sec:exp}
In the following subsections, we introduce datasets and evaluation metrics. Then we evaluate the performance of passenger prediction and passenger prediction based recommendation respectively.
\subsection{Dataset}
In our experiments, we use a real flight ticket dataset, consisting of all submitted orders for two years. An order contains account id and all passengers' id. We use desensitized passenger id to distinguish individuals from each other. In our experiment, we select active users whose amount of historical orders reaches a threshold, thus passengers' preferences can be extracted with more confidence. In addition, many researches\cite{amy:guess}\cite{yutaka:modeling} on shared account recommendation generate an artificial dataset by composing single accounts' data. We also generate an artificial dataset combining single passengers' data from two different accounts. The overviews of these two datasets and the meta data of orders are summarized in table \ref{tab:dat} and table \ref{tab:mdat}.
\begin{table}[!htbp]
\centering
\caption{\label{tab:dat}Datasets}
\begin{tabular}{|c|c|c|c|} \hline
& \# accounts & \# passengers & \# orders\\ \hline
Real & 4632&7034&38907 \\ \hline
Artificial &1604&3208&29759\\ \hline
\end{tabular}
\end{table}
